### 2012-10-07 sermons/Revelation1:4b.tex
.:wq

\subsection{His perfection}
\I  Note first that the sight of the Spirit as seven or seven-fold is
    \emph{not} a mistake.
\BI \I  The Spirit did not make a mistake in showing Himself to John
        as being seven-fold.
        In fact, he intentionally revealed Himself to John as such multiple times.
    \I  Moreover, John understood its importance and intentionally mentions
        the seven-fold Spirit at the very very outset of the letter.
    \I  Therefore, it must be important for us to understand what
        it means that the Spirit is seven-fold. \EI
\I  We can understand the significance of the seven-fold Spirit
    by observing the Biblical pattern 
    in which God uses a seven-fold repetition or progression to 
    signify completion or perfection.
\BI \I  Psalm 12:6, the perfection of God's word.
        \Q{ The words of the Lord are pure words, \\
            Like silver tried in a furnace of earth, \\
            Purified seven times.}
    \I  Genesis 1, the perfect, finished work of creation in seven days.
    \I  Leviticus 4:6, the finished atonement for a particular sin
        (of the high priest in this case),
        \Q{ The priest shall dip his finger in the blood and sprinkle some 
            of the blood seven times before the Lord, in front of the veil 
            of the sanctuary} \EI
\I  Of course, just because something is repeated or progresses in seven parts
    does not automatically make it perfect.
\BI \I  Matthew 18:21
        \Q{ Then Peter came to Him and said,
            ``Lord, how often shall my brother sin against me,
            and I forgive him? 
            Up to seven times?''} 
    \I  Rather, it is used by God to signify that something is perfect or complete. 
    \I  God sometimes uses a seven-fold repetition or progression
        to communicate perfection or completion to us. \EI
\I  It should be no surprise, then, that God working in a seven-fold manner
    abounds in Revelation, since it is a book of consummation.
\BI \I  This is most prominent in the three sets (seals, trumpets, bowls)
        of seven-fold judgements of God upon wicked humanity.
    \I  Revelation 15:1
        \Q{ Then I saw another sign in heaven, great and marvelous;
            seven angels having seven last plagues,
            for in them the wrath of God is complete.} \EI
\I  From this pattern, we understand that the seven-fold nature of the Spirit 
    communicates His perfection and completeness.
\BI \I  Not in any way lacking in any attribute of God.
    \I  He posesses all of the attributes of the Godhead to the fullest extent.
    \I  He is God undiminished and unabridged. 
        Nothing of God is toned down or omitted in Him.
    \I  He is an eternal, infinite being since 
        to be completely and perfectly divine is to be infinite.
    \I  He is all-knowing, all-powerful, all-capable.
    \I  He is holy, loving, and sovereign.
    \I  There is nothing He does not know, or does not control.
    \I  He is the complete infinity of Almighty God. \EI
\I  That He has a seven-fold nature means that
\BI \I  He is infinite in glory, and
    \I  that He is deserving of an eternity of worship. \EI
\I  It also means that if we were to study the Holy Spirit, 
    there would always be more to learn, and experience of Him.
\I  Show that He is all-capable in the salvation and sanctification of the believer.

\subsection{His submission}
\I  The Bible shows that the seven-fold Spirit of God is equal in person to 
    the Father and Son;
    however, it does not necessarily mean that He is equal in position, or role.
\I  From Revelation, we learn that He is lesser in authority than the Father.
\BI \I  From 1:4 and 4:5 we see the Father as being on the throne,
        and the Spirit as being in front of it.
    \I  Of course, the One on the throne has the supreme authority. \EI
\I  We also learn that He is under the authority of the Son.
\BI \I  Revelation 5:6
        \Q{ And I looked, and behold, in the midst of the throne and 
            of the four living creatues,
            and in the midst of the elders,
            stood a Lamb as though it had been slain,
            having seven horns and seven eyes,
            which are the seven Spirits of God sent out into all the earth.}
        The Spirit is pictured as belonging to the Lamb.
    \I  Revelation 3:1 is explicit.
        \Q{ These things says He who has the seven Spirits of God and 
            the seven stars: \ldots} \EI
\I  While equal in person, the Spirit joyfully submits in position/authority
    to both the Father and the Son.
\BI \I  No explicit mention in any worship scene in Revelation.
    \I  Does not appear on a throne, as the Father and Son do.
    \I  Instead, he directs worship to the Father and Son,
        as Galatians 4:6 says,
        \Q{ And because you are sons, God has sent forth the Spirit of His Son
            into your hearts, crying out, ``Abba, Father!''}
        The Holy Spirit does not direct us to worship Him, but the Father. \EI

@@

\section{Application -- How it matters}
\subsection{The Spirit indwells the believer}
\subsection{Be filled with the Spirit}
\EI

>>>>

\subsection{His perfection}
\I  Note first that the sight of the Spirit as seven or seven-fold is
    \emph{not} a mistake.
\BI \I  The Spirit did not make a mistake in showing Himself to John
        as being seven-fold.
        In fact, he intentionally revealed Himself to John as such multiple times.
    \I  Moreover, John understood its importance and intentionally mentions
        the seven-fold Spirit at the very very outset of the letter.
    \I  Therefore, it must be important for us to understand what
        it means that the Spirit is seven-fold. \EI
\I  We can understand the significance of the seven-fold Spirit
    by observing the Biblical pattern 
    in which God uses a seven-fold repetition or progression to 
    signify completion or perfection.
\BI \I  Psalm 12:6, the perfection of God's word.
        \Q{ The words of the Lord are pure words, \\
            Like silver tried in a furnace of earth, \\
            Purified seven times.}
    \I  Genesis 1, the perfect, finished work of creation in seven days.
    \I  Leviticus 4:6, the finished atonement for a particular sin
        (of the high priest in this case),
        \Q{ The priest shall dip his finger in the blood and sprinkle some 
            of the blood seven times before the Lord, in front of the veil 
            of the sanctuary} \EI
\I  Of course, just because something is repeated or progresses in seven parts
    does not automatically make it perfect.
\BI \I  Matthew 18:21
        \Q{ Then Peter came to Him and said,
            ``Lord, how often shall my brother sin against me,
            and I forgive him? 
            Up to seven times?''} 
    \I  Rather, it is used by God to signify that something is perfect or complete. 
    \I  God sometimes uses a seven-fold repetition or progression
        to communicate perfection or completion to us. \EI
\I  It should be no surprise, then, that God working in a seven-fold manner
    abounds in Revelation, since it is a book of consummation.
\BI \I  This is most prominent in the three sets (seals, trumpets, bowls)
        of seven-fold judgements of God upon wicked humanity.
    \I  Revelation 15:1
        \Q{ Then I saw another sign in heaven, great and marvelous;
            seven angels having seven last plagues,
            for in them the wrath of God is complete.} \EI
\I  From this pattern, we understand that the seven-fold nature of the Spirit 
    communicates His perfection and completeness.
\BI \I  Not in any way lacking in any attribute of God.
    \I  He posesses all of the attributes of the Godhead to the fullest extent.
    \I  He is God undiminished and unabridged. 
        Nothing of God is toned down or omitted in Him.
    \I  He is an eternal, infinite being since 
        to be completely and perfectly divine is to be infinite.
    \I  He is all-knowing, all-powerful, all-capable.
    \I  He is holy, loving, and sovereign.
    \I  There is nothing He does not know, or does not control.
    \I  He is the complete infinity of Almighty God. \EI
\I  That He has a seven-fold nature means that
\BI \I  He is infinite in glory, and
    \I  that He is deserving of an eternity of worship. \EI
\I  It also means that if we were to study the Holy Spirit, 
    there would always be more to learn, and experience of Him.
\I  His perfection is manifest in that He provides virtually all that
    we need as believers
\BI \I  Titus 3:5, by the Spirit we experience the new birth (regeneration).
        \Q{ \ldots but according to His mercy He saved us,
            through the washing of regeneration and renewing of the Holy Spirit,
            \ldots}
    \I  Romans 8:16, the Spirit gives us assurance of our salvation.
        \Q{The Spirit Himself bears witness with our spirit that we are children
            of God, \ldots}
    \I  Romans 8:13, by the Spirit we have victory of sin.
        \Q{ \ldots but if by the Spirit you put to death the deeds of the body,
            you will live}
    \I  Galatians 5:22, the Spirit sanctifies us.
        \Q{ But the fruit of the Spirit is love, joy, peace, longsuffering,
            kindness, goodness, faithfulness, gentleness, self-control.}
    \I  Romans 8:26, the Spirit intercedes for us.
        \Q{ \ldots the Spirit Himself makes intercession for us with 
            groanings which cannot be uttered.} \EI

\subsection{His submission}
\I  The Bible shows that the seven-fold Spirit of God is equal in person to 
    the Father and Son;
    however, it does not necessarily mean that He is equal in position, or role.
\I  From Revelation, we learn that He is lesser in authority than the Father.
\BI \I  From 1:4 and 4:5 we see the Father as being on the throne,
        and the Spirit as being in front of it.
    \I  Of course, the One on the throne has the supreme authority. \EI
\I  We also learn that He is under the authority of the Son.
\BI \I  Revelation 5:6
        \Q{ And I looked, and behold, in the midst of the throne and 
            of the four living creatues,
            and in the midst of the elders,
            stood a Lamb as though it had been slain,
            having seven horns and seven eyes,
            which are the seven Spirits of God sent out into all the earth.}
        The Spirit is pictured as belonging to the Lamb.
    \I  Revelation 3:1 is explicit.
        \Q{ These things says He who has the seven Spirits of God and 
            the seven stars: \ldots} \EI
\I  While equal in person, the Spirit joyfully submits in position/authority
    to both the Father and the Son.
\BI \I  No explicit mention in any worship scene in Revelation.
    \I  Does not appear on a throne, as the Father and Son do.
    \I  Instead, he directs worship to the Father and Son,
        as Galatians 4:6 says,
        \Q{ And because you are sons, God has sent forth the Spirit of His Son
            into your hearts, crying out, ``Abba, Father!''}
        The Holy Spirit does not direct us to worship Himself, but the Father. \EI

@@

\section{Application -- How it matters}
\subsection{The Spirit indwells the believer}
\I  The Spirit indwells us, Colossian 1:27
    \Q{To them God willed to make known what are the riches of the glory
        of this mystery among the Gentiles: which is Christ in you,
        the hope of glory.}
\I  What does this mean
\BI \I  Not physically, He is spirit after all.
    \I  It is that the Spirit interacts and works through our
        thoughts, attitudes and emotions.
        He works at the core of our being.
    \I  God is not simply making all things work together for our good
        through external means.
    \I  But the Spirit is directing our hearts and shaping our minds 
        to bring us comfort and strengthen us within. \EI
\I  Our greatest comfort and strength
\BI \I  It assures us of His love for us, Ephesians 1:13-14
        \Q{ In Him you also trusted, after you heard the word of truth,
            the gospel of our salvation; in whom also,
            having believed, you were sealed with the Holy Spirit of promise,
            who is a guarantee of our inheritance until the redemption
            of the purchased possession, to the praise of His glory.}
        We will never be vanquished by an outside foe.
    \I  It assures us that  \EI

\subsection{Be filled with the Spirit}
\I  Ephesians 4:18 
    \Q{ And do not be drunk with wine, in which is dissipation;
        but be filled with the Spirit.}
\EI
